Barneby Lives usage example

### 2023-05-18 scripts/project_template.Rmd
wrapper around parzer fleshed out

```{r dummy data}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:31, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  x = runif(1000, min = -120, max = -100), 
  y = runif(1000, min = 35, max = 48)
) %>% sf::st_as_sf(coords = c(x = 'x', y = 'y'), crs = 4326)

@@

 first convert from american date format (m%d%year%) to ISO 8601 for both collection and determination date

 second from ISO 8601 prepare notation dates for day of collection and date of determination in text format

```{r Date parsing}

#' create herbarium format dates for specimens
#' 
#' this function will return up to two additional columns with x. 'coll_date_text' a text format for the date of collection and 'det_date_text" a shorter text format date for identification
#' @param x an (sf/tibble/) data frame with both the collection date
#' and determination date
#' @param coll_date date of collection, expected to be of the format 'MM/DD/YYYY', minor checks for compliance to the format will be carried out before returning an error.
#' @param det_date date of determination, same format and processes as above. 
#' @returns coll_date ISO 8601, det_date ISO 8601, month, day, year, coll_date_text, det_date_text columns for printing onto the label
date_parser <- function(x, coll_date, det_date){

  coll_date_q <- enquo(coll_date)
  det_date_q <- enquo(det_date)

  names_v <- c('_dmy', '_day', '_mo',  '_yr', '_text')
  column_names <- c(paste0(coll_date, names_v), paste0(det_date, names_v) )

  mdy2dmy <- function(x){format(as.Date(x, format = "%m/%d/%Y"), "%d/%m/%Y")}

   x_dmy <- x |> 
    mutate(across(.cols = c(!!coll_date_q, !!det_date_q), lubridate::mdy, .names = "{.col}_dmy")) |>
    mutate(across(ends_with('_dmy'), ~ lubridate::month(.), .names = "{.col}_mo"),
           across(ends_with('_dmy'), ~ lubridate::day(.), .names = "{.col}_day"),
           across(ends_with('_dmy'), ~ lubridate::year(.), .names = "{.col}_yr")) |>
    mutate(across(.cols = c(!!coll_date_q, !!det_date_q), date2text, .names = '{.col}_text')) |>
    rename_with( ~ stringr::str_remove(., '_dmy'), matches("_dmy_.*$")) |>
    relocate(any_of(column_names), .before = last_col())

  return(x_dmy)
}

@@

```{r}

@@

browseVignettes("ggtree")
```

>>>>

```{r dummy data}
dummy_pts <- data.frame(
  collectionNo = paste0(rep('RCB', 1000), seq(1:1000)),
  day = c(sample(1:31, size = 1000, replace = T)),
  month = c(sample(5:10, size = 1000, replace = T)),
  year = 2023,
  longitude = runif(1000, min = -120, max = -100), 
  latitude = runif(1000, min = 35, max = 48)
) %>% sf::st_as_sf(coords = c(x = 'longitude', y = 'latitude'), crs = 4326, remove = F)

@@

```{r coordinate parsing if required}

library(parzer)

#' this function parses coordinates from DMS to decimal degrees
#'
#'@param x an input data frame to apply operations too
#'@param lat a name of the column holding the latitude values
#'@param long a name of the colymn holding the longitude values
#'@param dms are coordinates in degrees minutes seconds? TRUE for yes, FALSE for decimal degrees
dms2dd <- function(x, lat, long, dms){

  if(missing(lat)){
    lat = colnames(dummy_pts)[grep('lat', colnames(dummy_pts))] }
  if(missing(long)){
    long = colnames(dummy_pts)[grep('long', colnames(dummy_pts))] }

  if(missing(dms)){
    dms = long == parzer::parse_lon(long)
  }

  if(dms == T){
    x$latitude_dd = parzer::parse_lat(lat)
    x$longitude_dd = parzer::pase_lon(long)
  } else{
    colnames(x)[which(names(x) == lat)] <- 'latitude_dd'
    colnames(x)[which(names(x) == long)] <- 'latitude_dd'
  }

#  x[,c('latitude_dd', 'longitude_dd')] <- 
#    sapply(x[c('latitude_dd', 'latitude_dd')], as.numeric)

 # if(x$longitude_dd > 0){longitude_dd * -1}; 
#  if(x$latitude_dd < 0){latitude_dd * -1}

  # now overwrite the original DMS values in our exact format

  return(x)

  latitude_dms = paste0(
    'N ', parzer::pz_degree(x$latitude_dd),
    '°', round(parzer::pz_minute(x$latitude_dd), 2),
    "'", round(parzer::pz_second(x$latitude_dd), 2)
  )

  longitude_dms = paste0(
    'W ', parzer::pz_degree(x$longitude_dd), 
    '°', round(parzer::pz_minute(x$longitude_dd), 2), 
    "'", round(parzer::pz_second(x$longitude_dd), 2)
  )

}

out <- dms2dd(dummy_pts)

# str detect the lat column
# str detect the long column
# ensure input lat is North
# ensure input long is West
# str detect the datum column, else assume WGS84

@@

```{r date parsing}

@@
